### 2023-07-24 supplement/bibliography/rebib.tex

% !TeX root = RJwrapper.tex
\title{Working with bibliography in texor}
\author{by Abhishek Ulayil}

\maketitle

\abstract{
This is a small sample article to demonstrate usage of \CRANpkg{rebib} to manage bibliographies.
}

\section{Introduction}
The \CRANpkg{rebib} package is a spun-off from \CRANpkg{texor} package.

The decision to do this is the fact that the bibliography section 
in \CRANpkg{texor} package was expanding significantly, enough to deserve a dedicated space.

Thus \CRANpkg{rebib} was born and has improved over the past few months to include more advanced features and utilities.

\subsection{What does rebib do?}

In LaTeX, we have multiple ways to declare bibliography entries, which include embedding bibliography entries in the document itself or isolating that section to an external file commonly called a bbl file. Then we have the BibTeX format, which has its structure and is much superior in managing bibliographic entries.

LaTeX has no problem dealing with multiple formats and works well with all these bibliographic formats. However, some web publishing journals like RJ web articles from RJournal do not support embedded bibliographies. As a result, they need a BibTeX equivalent of the same.

The \CRANpkg{rebib} package converts embedded LaTeX bibliographies into a close  BibTeX equivalent. This package eases the author from the manual conversion of the documents and aids the texor project in automating the conversion of past RJournal articles to the RJ-web-article format.

\subsection{features of rebib}

\begin{itemize}
  \item Reads bib chunks to produce a very close BibTeX equivalent
  \item Title and author are usually mandatory fields
  \item URL, ISBN, publisher, pages and year are optional fields and will be enabled when relevant
  \item Rest of the data is stored in \verb|"journal"|(internally) and \verb|"publisher"|(when writing BibTeX file)
  \item Ignores commented LaTeX code 
  \item Citation tracker
  \item Bibliography Aggregation
\end{itemize}

\section{Parser}

The magic behind \CRANpkg{rebib} is a parser which parses bibliographic data and assorts them according to the matching regular expressions.

\subsubsection{Stage 1: Read the embedded bibliography block}

This stage is a minor step where it reads the Embedded Bibliography from the LaTeX document. This step also includes Filtering out the commented code to avoid un-intended entries read.

Lastly, the data is broken down based on the LaTeX macro \verb|\\bibitem| as a marker for a new entry and this assorted data is exported to a variable.

\begin{Sinput}
dir.create(your_article_folder <- file.path(tempdir(), "exampledir"))
example_files <-  system.file("article", package = "rebib")
x <- file.copy(from = example_files,to=your_article_folder,recursive = T)
your_article_path <- paste(your_article_folder,"article",sep="/")
file_name <- rebib:::get_texfile_name(your_article_path)
bib_items <- rebib:::extract_embeded_bib_items(your_article_path,file_name)
bib_items[[1]]
#> [1] "\\bibitem[Ihaka, Ross and Gentleman, Robert]{ihaka:1996}"                          
#> [2] "Ihaka, Ross and Gentleman, Robert"                                                 
#> [3] "\\newblock \\emph{R: A Language for Data Analysis and Graphics.}"                  
#> [4] "\\newblock \\emph{Journal of Computational and Graphical Statistics}, 3:\\penalty0"
#> [5] "299--314, 1996."                                                                   
#> [6] "\\newblock URL : \\url{https://doi.org/10.1080/10618600.1996.10474713}"
\end{Sinput}

\subsubsection{Stage 2: Regex powered parser}

Now, with the chunks of bibliographic entries, each is passed to a parser which will break it down based on regular expressions.
The logic is to use the LaTeX macro \verb|\\newblock| as a placeholder to identify the position of text elements relative to it.

The first value to be parsed is the \verb|unique_id| also called the citation reference which is used to cite elements inside the article. Usually, this is in the first or second line of the whole entry. The position of the 
\verb|unique_id| will determine the position of the author names.

\begin{Soutput}
bib_items[[1]][1]
#> [1] "\\bibitem[Ihaka, Ross and Gentleman, Robert]{ihaka:1996}"
\end{Soutput}

After reading the \verb|unique_id|, the parser will attempt to read the author name(s).

\begin{Sinput}
bib_items[[1]][2]
#> [1] "Ihaka, Ross and Gentleman, Robert"
\end{Sinput}

Next, the title is extracted based on the position of the new blocks or the end of the bib chunk.

\begin{Sinput}
bib_items[[1]][3]
#> [1] "\\newblock \\emph{R: A Language for Data Analysis and Graphics.}"
\end{Sinput}

This way the crucial elements of the bibliographic entry (unique\_id, author names and title ) are parsed out.
The remaining data is stored as \verb|journal| internally and \verb|publisher| when writing to a new BibTeX file.

\begin{Sinput}
bib_items[[1]][4:6]
#> [1] "\\newblock \\emph{Journal of Computational and Graphical Statistics}, 3:\\penalty0"
#> [2] "299--314, 1996."                                                                   
#> [3] "\\newblock URL : \\url{https://doi.org/10.1080/10618600.1996.10474713}"
\end{Sinput}

A series of filters for ISBN, URL, pages and year fields are applied to search for relevant data from the remaining data. If the data is not available then it is set as NULL and will be skipped while writing the BibTeX file. 
There is a lot of filtering of common LaTeX elements and after that, the data remaining is stored in a structured format to be written to a file.

\begin{Schunk}
\begin{Soutput}
bib_entry <- rebib:::bib_handler(bib_items)
bib_entry
#> $book
#> $book[[1]]
#> $book[[1]]$unique_id
#> [1] "ihaka:1996"
#> 
#> $book[[1]]$author
#> [1] "Ihaka, Ross and Gentleman, Robert"
#> 
#> $book[[1]]$title
#> [1] "R: A Language for Data Analysis and Graphics"
#> 
#> $book[[1]]$journal
#> [1] "Journal of Computational and Graphical Statistics 3:   :"
#> 
#> $book[[1]]$year
#> [1] "1996"
#> 
#> $book[[1]]$URL
#> [1] "https://doi.org/10.1080/10618600.1996.10474713"
#> 
#> $book[[1]]$isbn
#> NULL
#> 
#> $book[[1]]$pages
#> [1] "299--314"
#> 
#> 
#> $book[[2]]
#> $book[[2]]$unique_id
#> [1] "R"
#> 
#> $book[[2]]$author
#> [1] "R Core Team"
#> 
#> $book[[2]]$title
#> [1] "R: A Language and Environment for Statistical Computing"
#> 
#> $book[[2]]$journal
#> [1] "R Foundation for Statistical Computing Vienna Austria    :"
#> 
#> $book[[2]]$year
#> [1] "2016"
#> 
#> $book[[2]]$URL
#> [1] "https://www.R-project.org/"
#> 
#> $book[[2]]$isbn
#> [1] "3-900051-07-0"
#> 
#> $book[[2]]$pages
#> NULL
\end{Soutput}
\end{Schunk}

\subsubsection{Stage 3: BibTeX writer}

After reading the bibliographic entries and splitting out meaningful values from them, we can finally write a structured file in the BibTeX format.

The writer will read the bib chunks one at a time based on the metadata extracted and will write the corresponding data fields. The default entry type is a book, which should not have any problems with the web articles. 
\begin{Schunk}

\begin{Sinput}
file_path <- paste0(your_article_path,"/",file_name)
bib_path <- paste0(your_article_path,"/example.bib")
rebib:::bibtex_writer(bib_entry,file_path)
cat(readLines(paste(your_article_path,"example.bib",sep="/")),sep = "\n")
#> @book{ihaka:1996,
#> author = {{Ihaka, Ross and Gentleman, Robert}},
#> title = {{R: A Language for Data Analysis and Graphics}},
#> publisher = {Journal of Computational and Graphical Statistics 3:   :},
#> pages = {299--314},
#> year = {1996},
#> url = {https://doi.org/10.1080/10618600.1996.10474713}
#> }
#> @book{R,
#> author = {R {Core Team}},
#> title = {{R: A Language and Environment for Statistical Computing}},
#> publisher = {R Foundation for Statistical Computing Vienna Austria    :},
#> year = {2016},
#> url = {https://www.R-project.org/},
#> isbn = {3-900051-07-0}
#> }
\end{Sinput}
\end{Schunk}

I expect the authors who are converting the document to take a look and check if there are any errors or missing values.

\section{General usage}
In general to use \CRANpkg{rebib} package, it is quite easy with requiring only one function call.
\subsubsection{Using with RJournal article}
In bibliography construction mode, \CRANpkg{rebib} will check for existing BibTeX files and will convert the embedded bibliography only when there are no linked BibTeX files in the RJ article.
\begin{verbatim}
%% typically rebib will search for this line in the RJ article, 
%% not the RJwrapper file
\bibliography{example}
\end{verbatim}
If there is no linked BibTeX file, \CRANpkg{rebib} will begin its procedure of converting the embedded bibliography to BibTeX and then link it with the article file as well.
\begin{Schunk}
\begin{Sinput}
# for files without BibTeX source
rebib::handle_bibliography(your_article_path)
#> Warning in get_bib_file(article_dir, file_name): No Bib files found !
#> BibTeX file does not exist
#> will parse for bibliography
#> bibtex file created
cat(readLines(paste(your_article_path,"example.bib",sep="/")),sep = "\n")
#> @book{ihaka:1996,
#> author = {{Ihaka, Ross and Gentleman, Robert}},
#> title = {{R: A Language for Data Analysis and Graphics}},
#> publisher = {Journal of Computational and Graphical Statistics 3:   :},
#> pages = {299--314},
#> year = {1996},
#> url = {https://doi.org/10.1080/10618600.1996.10474713}
#> }
#> @book{R,
#> author = {R {Core Team}},
#> title = {{R: A Language and Environment for Statistical Computing}},
#> publisher = {R Foundation for Statistical Computing Vienna Austria    :},
#> year = {2016},
#> url = {https://www.R-project.org/},
#> isbn = {3-900051-07-0}
#> }
\end{Sinput}
\end{Schunk}

\subsubsection{Bibliography aggregation}

If you have a BibTeX file and it is missing some references then this mode can help you read the embedded bibliography and combine the two bibliographies in a single BibTeX file.
\begin{verbatim}
# for files with BibTeX source as well as embedded entries
rebib::aggregate_bibliography(your_article_path)
#> Found Bib file  example.bib
#> bibliography aggregation possible
#> Found Bib file  example.bib
#> aggregation delta : 2
#> BibTeX file aggregated
\end{verbatim}

\subsubsection{Using with any LaTeX/BBL file}
The bibliography converter will convert any LaTex or BBL file given as a path. This way, one can generate BibTeX from the embedded bibliography or BBL file.
\begin{verbatim}
rebib::biblio_converter(path_to_your_file)
#> working directory : /temp/Rtmpt6Qk0v/exampledir/standalone
#> bib entries : 6
#> Written BibTeX file : /temp/Rtmpt6Qk0v/exampledir/standalone/sample.bib
\end{verbatim}

\section{Summary}

In summary the \CRANpkg{rebib} package supports:
\begin{itemize}
\item Conversion of embedded natbib type bibliography used in RJournal
\item Aggregation of bibliographic entries from the article as well as BibTeX
\end{itemize}

\address{%
Abhishek Ulayil\\
Student, Institute of Actuaries of India\\%
Mumbai, India\\
ORCiD: 0009-0000-6935-8690\\
}