### 2023-08-15 supplement/figure-env/figure-example.tex
Updated figures and clarifications on differences in CodeBlocks

% !TeX root = RJwrapper.tex
\title{Working with figure environments in texor}
\author{by Abhishek Ulayil}

\maketitle

\abstract{
This is a small sample article to demonstrate usage of texor to convert figure environments.
}

\section{Introduction}

Images are an essential component in any article, However due to the differences in the support for various graphic formats between LaTeX and markdown/HTML we need to fallback on raster graphics. 
It is well summarized in the Table \ref{table:1}.

\begin{table}[htbp]
\centering
\begin{tabular}{l | llll }
 \hline
 Graphics Format & LaTeX & Markdown & Rmarkdown & HTML \\
 \hline
 PNG       & Yes & Yes & Yes & Yes \\
 JPG       & Yes & Yes & Yes & Yes \\
 PDF       & Yes & No & No & No \\
 SVG       & No & Yes & Yes & Yes \\
 Tikz      & Yes & No & Yes & No \\
 Algorithm & Yes & No & No & No \\
\hline
\end{tabular}
\caption{Image Format support in various Markup/Typesetting Languages}
\label{table:1}
\end{table}

\section{Image with width parameters}
This section may contain a figure such as Figure~\ref{figure:rlogo}.
This is the most basic example of figure.

\begin{verbatim}
\begin{figure}[htbp]
  \centering
  \includegraphics[width=0.35\textwidth]{Rlogo-5.png}
  \caption{The logo of R.}
  \label{figure:rlogo}
\end{figure}
\end{verbatim}

\begin{figure}[htbp]
  \centering
  \includegraphics[width=0.35\textwidth]{Rlogo-5.png}
  \caption{The logo of R.}
  \label{figure:rlogo}
\end{figure}

\section{Images in PDF format}

Image \ref{fig:normal} is a graphical representation of normal distribution.

\begin{verbatim}
\begin{figure}[htbp]
  \centering
  \includegraphics[width=0.5\textwidth]{normal}
  \caption{PDF of a normal distribution}
  \label{fig:normal}
\end{figure}
\end{verbatim}
\begin{figure}[htbp]
  \centering
  \includegraphics[width=0.5\textwidth]{normal}
  \caption{PDF of a normal distribution}
  \label{fig:normal}
\end{figure}

\section{Multiple images}
 Pandoc v3 and above now supports a new Figure object \citep{pandoc} which supports multiple 
 images side by side or in a grid format.
%% Two images side by side
\subsubsection{Two or more Images side by side}
\begin{verbatim}
\begin{figure}[htbp]
  \centering
  \includegraphics[width=0.45\textwidth]{Rlogo-5.png}\includegraphics[width=0.45\textwidth]{normal}
  \caption{Images side by side}
  \label{fig:twoimages}
\end{figure}
\end{verbatim}

\begin{figure*}[htbp]
  \centering
  \includegraphics[width=0.45\textwidth]{Rlogo-5.png}\includegraphics[width=0.45\textwidth]{normal}
  \caption{Images side by side}
  \label{fig:twoimages}
\end{figure*}

%% Four images in a grid
\subsubsection{Four Images in a grid}

\begin{verbatim}
\begin{figure}[htbp]
  \centering
  \includegraphics[width=0.45\textwidth]{Rlogo-5.png}\includegraphics[width=0.45\textwidth]{normal}
  \includegraphics[width=0.45\textwidth]{normal}\includegraphics[width=0.45\textwidth]{Rlogo-5.png}
  \caption{Multiple images in a grid}
  \label{fig:fourimages}
\end{figure}

\end{verbatim}

\begin{figure}[htbp]
  \centering
  \includegraphics[width=0.45\textwidth]{Rlogo-5.png}\includegraphics[width=0.45\textwidth]{normal}
  \includegraphics[width=0.45\textwidth]{normal}\includegraphics[width=0.45\textwidth]{Rlogo-5.png}
  \caption{Multiple images in a grid}
  \label{fig:fourimages}
\end{figure}

\section{Tikz images}

Here is a tikz image example in fig \ref{fig:tikz} adapted from \citep{casflow}.

Another interesting aspect of including tikz image here is that you can modify the
source and re-convert without making any other changes. Figure will get updated
in the generated html article.

\subsubsection{Tikz Code:}
The figure  \ref{fig:tikz} is an example of how \pkg{texor} handles tikz images,

\begin{itemize}
\item First figures containing tikz images are isolated.
\item Tikz libraries are fetched from the wrapper file,
\item Tikz Code section is isolated into a standalone LaTeX file and compiled.
\item The Compiled LaTeX file generates a PDF Image.
\item This is then Converted again to PNG format as PDFs embedded in HTML wont look good.
\item In the figure environment a line is added \verb|\includegraphics{tikz/somefigure.png}|
\item During the conversion in pandoc, a lua filter removes extra/unncessary text from the rest of the figure
environemnt.
\end{itemize}

If You use \pkg{texor} to convert your articles using \verb|texor::latex_to_web()| with \verb|temp_mode=TRUE|(it is TRUE by default). The resultant Rmarkdown/HTML file will not modify the contents of your LaTeX file. In this case you can keep reloading the article after making changes to the tikz images, without having to do the above steps manually in case you are converting the article by hand.

\begin{figure*}

%% Generated Image will included as a PNG above
  \centering
\tikzstyle{process} = [rectangle, rounded corners,
minimum width=3cm, 
minimum height=1cm,
text centered, 
draw=black]
\tikzstyle{arrow} = [thick,->,>=stealth]
\begin{tikzpicture}[node distance=2cm]
%Nodes
\node (start) [process] {Tikz};
\node (isolate) [process, below right of=start] {Isolated LaTeX};
\node (pro1) [process, below right of=isolate] {PDF};
\node (pro2) [process, below right of=pro1, yshift=-0.5cm] {PNG};
% arrows
\draw [arrow] (start) -- node[anchor=east] {Extract} (isolate);
\draw [arrow] (isolate) -- node[anchor=east] {Compile} (pro1);
\draw [arrow] (pro1) -- node[anchor=east] {Rasterize} (pro2);

\end{tikzpicture}
\caption{Tikz Image example}
  \label{fig:tikz}
\end{figure*}

\section{Algorithm2e diagrams}

we do support algorithm2e diagrams and images, these will be numbered differently
and we strongly suggest to use "alg:" in labels for best results. 
Here is an example of algorithm \ref{alg:how}, referenced from \cite{algoexample}

\begin{algorithm}[htbp]
\SetAlgoLined
\KwData{this text}
\KwResult{how to write algorithm with \LaTeX2e }
initialization\;
\While{not at end of this document}{
read current\;
\eIf{understand}{
go to next section\;
current section becomes this one\;
}{
go back to the beginning of current section\;
}
}
\caption{How to write algorithms}
  \label{alg:how}
\end{algorithm}

\section{Other elements in figure objects}

Figures can also house non-image environments like CodeBlocks and WideTables. This is 
the part where things start to look a bit different in the converted articles, CodeBlocks 
in figure environments would have numbering and caption that goes "Figure x:". However, 
when the same CodeBlock is converted to Rmarkdown, they are numbered as "CodeBlock x:".
This distinguishes them from normal Images and matches with authors vision more appropriately.

As for Inclusion of widetables, since it houses two different tabular environments with the same caption,
it would be treated as two different tables by pandoc sharing the same caption. This would not be the best solution hence, \pkg{texor} package uses Figure enviornment to house the widetables.
This way they are distinguished well from normal tables, also the numbering scheme is seperated from figures "WideTable x:".

However one important thing to note is, the reference links and reference numbering would be shared with ones of the figures. So the changes that the package makes are cosmetic to captions. Like CodeBlock 1 below refers to \ref{code:example}. The number should have been 1 but instead it is 7.

\begin{figure}[htbp]
\begin{center}
\begin{verbatim}
code_in_figure <- function() {
  if (pandoc_version >= 3) {
    print("Code in Figure Supported")
  }
  else {
    print("code in Figure not supported")
  }
}
\end{verbatim}
\caption{ Example Code inside Figure environment}
\label{code:example}
\end{center}
\end{figure}
\section{Summary}

In summary the \CRANpkg{texor} package supports:
\begin{itemize}
\item Almost all image formats in LaTeX.
\item Algorithm and tikz as well in some capacity.
\item Multiple images in grid,side-by-side configuration.
\item Image Captions with Numbering and Labelling.
\end{itemize}

\begin{thebibliography}{4}
    \providecommand{\natexlab}[1]{#1}
    \providecommand{\url}[1]{\texttt{#1}}
    \expandafter\ifx\csname urlstyle\endcsname\relax
      \providecommand{\doi}[1]{doi: #1}\else
      \providecommand{\doi}{doi: \begingroup \urlstyle{rm}\Url}\fi
\bibitem[Cassidy(2013)]{casflow}
Josh Cassidy
\newblock LaTeX Graphics using TikZ: A Tutorial for Beginners (Part 3)—Creating Flowcharts
\newblock \emph{Overleaf tutorials} \penalty0 2013
\newblock URL \url{https://www.overleaf.com/learn/latex/}

\bibitem[Fiorio (2017)]{algoexample}
Christophe Fiorio
\newblock algorithm2e.sty — package for algorithms, release 5.2
\newblock \emph{CTAN},\penalty0 2017
\newblock URL \url{https://mirror.kku.ac.th/CTAN/macros/latex/contrib/algorithm2e/doc/algorithm2e.pdf}

\bibitem[Krewinkel, Lucero (2023)]{pandoc}
Albert Krewinkel and Aner Lucero
\newblock pandoc 3.0 Release notes
\newblock \emph{pandoc} \penalty0 2023
\newblock URL \url{https://pandoc.org/releases.html}

\end{thebibliography}

\address{%
Abhishek Ulayil\\
Student, Institute of Actuaries of India\\%
Mumbai, India\\
ORCiD: 0009-0000-6935-8690\\
}