### 2023-08-21 supplement/figure-env/figure-example.tex
Updated supplementary to fix some #2 issues except math and lua filters

% !TeX root = RJwrapper.tex
\title{Working with figure environments in texor}
\author{by Abhishek Ulayil}

\maketitle

\abstract{
This is a small sample article to demonstrate usage of \CRANpkg{texor} to convert figure environments.
}

\section{Introduction}

Images are an essential component of any article, however, due to the differences in support for various graphic formats between LaTeX and markdown/HTML we need to fall back on raster graphics. 
The support for different image formats across markup languages is summarized in Table~\ref{table:1}.

\begin{table}[htbp]
\centering
\begin{tabular}{l | llll }
 \hline
 Graphics Format & LaTeX & Markdown & Rmarkdown & HTML \\
 \hline
 PNG       & Yes & Yes & Yes & Yes \\
 JPG       & Yes & Yes & Yes & Yes \\
 PDF       & Yes & No & No & No \\
 SVG       & No & Yes & Yes & Yes \\
 Tikz      & Yes & No & Yes & No \\
 Algorithm & Yes & No & No & No \\
\hline
\end{tabular}
\caption{Image Format support in various Markup/Typesetting Languages}
\label{table:1}
\end{table}

\section{Image with width parameters}

The following code includes an image with width parameters, producing the output in Figure~\ref{figure:rlogo}.

\begin{verbatim}
\begin{figure}[htbp]
  \centering
  \includegraphics[width=0.35\textwidth]{Rlogo-5.png}
  \caption{The logo of R.}
  \label{figure:rlogo}
\end{figure}
\end{verbatim}

\begin{figure}[htbp]
  \centering
  \includegraphics[width=0.35\textwidth]{Rlogo-5.png}
  \caption{The logo of R.}
  \label{figure:rlogo}
\end{figure}

This is the most basic example of figure. Similarly PDF figures will be included as rasterized images.

\section{Multiple images}
 Pandoc v3 and above now support a new \code{Figure} object \citep{pandoc} which supports multiple 
 images side by side or in a grid format. The following subsections demonstrate these capabilities.
%% Two images side by side
\subsubsection{Two or more Images side by side}
\begin{verbatim}
\begin{figure}[htbp]
  \centering
  \includegraphics[width=0.45\textwidth]{Rlogo-5.png}\includegraphics[width=0.45\textwidth]{normal}
  \caption{Images side by side}
  \label{fig:twoimages}
\end{figure}
\end{verbatim}

\begin{figure*}[htbp]
  \centering
  \includegraphics[width=0.45\textwidth]{Rlogo-5.png}\includegraphics[width=0.45\textwidth]{normal}
  \caption{Images side by side}
  \label{fig:twoimages}
\end{figure*}

%% Four images in a grid
\subsubsection{Four Images in a grid}

\begin{verbatim}
\begin{figure}[htbp]
  \centering
  \includegraphics[width=0.45\textwidth]{Rlogo-5.png}\includegraphics[width=0.45\textwidth]{normal}
  \includegraphics[width=0.45\textwidth]{normal}\includegraphics[width=0.45\textwidth]{Rlogo-5.png}
  \caption{Multiple images in a grid}
  \label{fig:fourimages}
\end{figure}

\end{verbatim}

\begin{figure}[htbp]
  \centering
  \includegraphics[width=0.45\textwidth]{Rlogo-5.png}\includegraphics[width=0.45\textwidth]{normal}
  \includegraphics[width=0.45\textwidth]{normal}\includegraphics[width=0.45\textwidth]{Rlogo-5.png}
  \caption{Multiple images in a grid}
  \label{fig:fourimages}
\end{figure}

\section{Tikz images}

The \pkg{texor} package supports tikz images by rasterizing it and making it web friendly.
Figure \ref{fig:tikz} shows a tikz image adapted from \citep{casflow}.

\subsubsection{Tikz Code:}
The image in Figure  \ref{fig:tikz} is a graphical representation of how \pkg{texor} handles tikz images:

\begin{itemize}
\item Figures containing tikz images are isolated.
\item Tikz libraries are fetched from the wrapper file.
\item The Tikz code section is isolated into a standalone LaTeX file and compiled.
\item The compiled LaTeX file generates a PDF Image.
\item This is converted to PNG format for embedding in the HTML output.
\item \verb|\includegraphics{tikz/somefigure.png}| is added to the figure environment.
\item A lua filter removes redundant text the figure environment.
\end{itemize}

If you use \pkg{texor} to convert your articles using \verb|texor::latex_to_web()| with \verb|temp_mode=TRUE|(it is \verb|TRUE| by default). The resultant Rmarkdown/HTML file will not modify the contents of your LaTeX file. In this case you can keep reloading the article after making changes to the tikz images, without having to do the above steps manually in case you are converting the article by hand.

\begin{figure*}

%% Generated Image will included as a PNG above
  \centering
\tikzstyle{process} = [rectangle, rounded corners,
minimum width=3cm, 
minimum height=1cm,
text centered, 
draw=black]
\tikzstyle{arrow} = [thick,->,>=stealth]
\begin{tikzpicture}[node distance=2cm]
%Nodes
\node (start) [process] {Tikz};
\node (isolate) [process, below right of=start] {Isolated LaTeX};
\node (pro1) [process, below right of=isolate] {PDF};
\node (pro2) [process, below right of=pro1, yshift=-0.5cm] {PNG};
% arrows
\draw [arrow] (start) -- node[anchor=east] {Extract} (isolate);
\draw [arrow] (isolate) -- node[anchor=east] {Compile} (pro1);
\draw [arrow] (pro1) -- node[anchor=east] {Rasterize} (pro2);

\end{tikzpicture}
\caption{Tikz Image example}
  \label{fig:tikz}
\end{figure*}

\section{Algorithm2e diagrams}

Diagrams and images using the \code{algorithm2e} environment are supported, these will be numbered differently.
We strongly suggest to use "alg:" in labels for best results. As a part of the filtering requires "alg:" to number
algorithm references seperately, otherwise they will share references with normal figures.

Algorithm \ref{alg:how} is an example from the \pkg{algorithm2e} vignette \cite{algoexample}.

\begin{algorithm}[htbp]
\SetAlgoLined
\KwData{this text}
\KwResult{how to write algorithm with \LaTeX2e }
initialization\;
\While{not at end of this document}{
read current\;
\eIf{understand}{
go to next section\;
current section becomes this one\;
}{
go back to the beginning of current section\;
}
}
\caption{How to write algorithms}
  \label{alg:how}
\end{algorithm}

\section{Other elements in figure objects}

Figures can also house non-image environments like code blocks. Code blocks 
in figure environments would share numbering with normal figures, similar to LaTeX.
Here Figure \ref{code:example} refers to a code block.

\begin{figure}[htbp]
\begin{center}
\begin{verbatim}
code_in_figure <- function() {
  if (pandoc_version >= 3) {
    print("code in figure supported")
  }
  else {
    print("code in figure not supported")
  }
}
\end{verbatim}
\caption{ Example Code inside Figure environment}
\label{code:example}
\end{center}
\end{figure}

\section{Limitations}

The limitations of \CRANpkg{texor} package in figure handling are:

\begin{itemize}
\item Animations using the \pkg{animation} package in LaTeX wont be supported. The authors
can include GiF equivalents of the same in the Rmarkdown/web article.
\item Multi-page PDF files are currently not supported.
\item Usage of subfigure,sclaebox,wrapfigure environments will not work as intended.
\item Currently only \verb|[width=X.X\textwdith]| format is supported for defining the width
in web article (default width would be 100\% in other cases).
\item Other \verb|\inclduegraphics[..]| options will be ignored.
\end{itemize}

\section{Summary}

In summary the \CRANpkg{texor} package supports:

\begin{itemize}
\item Multiple images in grid,side-by-side configuration.
\item Image Captions with Numbering and Labelling.
\item Algorithm2e and tikz figures.
\end{itemize}

\begin{thebibliography}{4}
    \providecommand{\natexlab}[1]{#1}
    \providecommand{\url}[1]{\texttt{#1}}
    \expandafter\ifx\csname urlstyle\endcsname\relax
      \providecommand{\doi}[1]{doi: #1}\else
      \providecommand{\doi}{doi: \begingroup \urlstyle{rm}\Url}\fi
\bibitem[Cassidy(2013)]{casflow}
Josh Cassidy
\newblock LaTeX Graphics using TikZ: A Tutorial for Beginners (Part 3)—Creating Flowcharts
\newblock \emph{Overleaf tutorials} \penalty0 2013
\newblock URL \url{https://www.overleaf.com/learn/latex/}

\bibitem[Fiorio (2017)]{algoexample}
Christophe Fiorio
\newblock algorithm2e.sty — package for algorithms, release 5.2
\newblock \emph{CTAN},\penalty0 2017
\newblock URL \url{https://mirror.kku.ac.th/CTAN/macros/latex/contrib/algorithm2e/doc/algorithm2e.pdf}

\bibitem[Krewinkel, Lucero (2023)]{pandoc}
Albert Krewinkel and Aner Lucero
\newblock pandoc 3.0 Release notes
\newblock \emph{pandoc} \penalty0 2023
\newblock URL \url{https://pandoc.org/releases.html}

\end{thebibliography}

\address{%
Abhishek Ulayil\\
Student, Institute of Actuaries of India\\%
Mumbai, India\\
ORCiD: 0009-0000-6935-8690\\
}