### 2025-09-07 supplement/bibliography/RJwrapper.Rmd
Final response for inclusion of Rmd files.

---
title: Handling bibliographies with texor
abstract: |
  This is a small sample article to demonstrate usage of
  [**rebib**](https://CRAN.R-project.org/package=rebib) to manage
  bibliographies.
author:
- name: Abhishek Ulayil
  affiliation: Student, Institute of Actuaries of India
  orcid: 0009-0000-6935-8690
  address: Mumbai, India
date: '2025-09-01'
date_received: ~
journal:
  firstpage: ~
  lastpage: ~
volume: 17
issue: 3
slug: RJ-2025-000
citation_url: https://rjournal.github.io/
packages:
  cran:
  - rebib
  - texor
  bioc: []
preview: preview.png
bibliography: ~
CTV: ~
legacy_pdf: yes
legacy_converted: yes
output:
  rjtools::rjournal_web_article:
    self_contained: yes
    toc: no
    mathjax: https://cdn.jsdelivr.net/npm/mathjax@3/es5/tex-chtml-full.js
    md_extension: -tex_math_single_backslash

---

::: article
## Introduction

The [**rebib**](https://CRAN.R-project.org/package=rebib) package is a
spin-off from the [**texor**](https://CRAN.R-project.org/package=texor)
package.

The spin-off cam about because the bibliography section in
[**texor**](https://CRAN.R-project.org/package=texor) package was
expanding significantly, enough to deserve a dedicated space.

Thus [**rebib**](https://CRAN.R-project.org/package=rebib) was born and
it has improved over the past few months to include more advanced
features and utilities.

### What does rebib do?

In LaTeX, we have multiple ways to declare bibliography entries, which
include embedding entries formatted to a given style in the document
itself or isolating that section to an external file commonly called a
`.bbl` file. Then we have the BibTeX format, which is designed to store
the full bibliographic information for each reference, making it
superior for managing references.

LaTeX has no problem dealing with multiple formats and works well with
all these bibliographic formats. However, some web publishing journals
like R Markdown articles for the R Journal do not support embedded
bibliographies. As a result, they need a BibTeX equivalent of the same.

The [**rebib**](https://CRAN.R-project.org/package=rebib) package
converts embedded LaTeX bibliographies into a close BibTeX equivalent.
This package relieves the author from manual conversion of the documents
and aids the project to convert legacy R Journal articles to the current
R Markdown format.

### Features of rebib

-   Extracts embedded bibliographic entries from a `.tex` file.

-   Creates the mandatory title and author fields.

-   Creates the optional URL, ISBN, publisher, pages and year fields,
    when available.

-   Stores the remaining information in `"journal"` (internally) and
    `"publisher"` (when writing BibTeX file).

-   Ignores commented LaTeX code.

-   Tracks citations included in the document.

-   Aggregates embedded bibliographic entries with references from
    supplementary `.bbl` and `.bib` files.

## Parser

The magic behind [**rebib**](https://CRAN.R-project.org/package=rebib)
is a parser which parses bibliographic data and sorts them according to
the matching regular expressions.

#### Stage 1: Read the embedded bibliography block

This stage is a minor step where it reads the Embedded Bibliography from
the LaTeX document. This step also includes filtering out commented
lines to avoid reading un-intended entries.

The bibliography is split into entries using the LaTeX macro `\\bibitem`
as a marker for a new entry and the resulting data is exported to a
variable.

``` r
dir.create(your_article_folder <- file.path(tempdir(), "exampledir"))
example_files <-  system.file("article", package = "rebib")
x <- file.copy(from = example_files,to=your_article_folder,recursive = T)
your_article_path <- paste(your_article_folder,"article",sep="/")
file_name <- rebib:::get_texfile_name(your_article_path)
bib_items <- rebib:::extract_embeded_bib_items(your_article_path,file_name)
bib_items[[1]]
#> [1] "\\bibitem[Ihaka, Ross and Gentleman, Robert]{ihaka:1996}"                          
#> [2] "Ihaka, Ross and Gentleman, Robert"                                                 
#> [3] "\\newblock \\emph{R: A Language for Data Analysis and Graphics.}"                  
#> [4] "\\newblock \\emph{Journal of Computational and Graphical Statistics}, 3:\\penalty0"
#> [5] "299--314, 1996."                                                                   
#> [6] "\\newblock URL : \\url{https://doi.org/10.1080/10618600.1996.10474713}"
```

#### Stage 2: Regex powered parser

Now, with the chunks of bibliographic entries, each is passed to a
parser which will break it down based on regular expressions. The logic
is to use the LaTeX macro `\\newblock` as a placeholder to identify the
position of text elements relative to it.

The first value to be parsed is the `unique_id`, also called the
citation reference which is used to cite elements inside the article.
Usually, this is in the first or second line of the whole entry. The
position of the `unique_id` will determine the position of the author
names.

``` r
bib_items[[1]][1]
#> [1] "\\bibitem[Ihaka, Ross and Gentleman, Robert]{ihaka:1996}"
```

After reading the `unique_id`, the parser will attempt to read the
author name(s).

``` r
bib_items[[1]][2]
#> [1] "Ihaka, Ross and Gentleman, Robert"
```

Next, the title is extracted based on the position of the new blocks or
the end of the bib chunk.

``` r
bib_items[[1]][3]
#> [1] "\\newblock \\emph{R: A Language for Data Analysis and Graphics.}"
```

This way the crucial elements of the bibliographic entry (unique_id,
author names and title) are parsed out. The remaining data is stored as
`journal` internally and `publisher` when writing to a new BibTeX file.

``` r
bib_items[[1]][4:6]
#> [1] "\\newblock \\emph{Journal of Computational and Graphical Statistics}, 3:\\penalty0"
#> [2] "299--314, 1996."                                                                   
#> [3] "\\newblock URL : \\url{https://doi.org/10.1080/10618600.1996.10474713}"
```

A series of filters for ISBN, URL, pages and year fields are applied to
search for relevant data from the remaining data. If the data is not
available, then it is set as `NULL` and will be skipped while writing
the BibTeX file. There is a lot of filtering of common LaTeX elements
and after that, the remaining data is stored in a structured format to
be written to a file.

``` r
bib_entry <- rebib:::bib_handler(bib_items)
bib_entry
#> $book
#> $book[[1]]
#> $book[[1]]$unique_id
#> [1] "ihaka:1996"
#> 
#> $book[[1]]$author
#> [1] "Ihaka, Ross and Gentleman, Robert"
#> 
#> $book[[1]]$title
#> [1] "R: A Language for Data Analysis and Graphics"
#> 
#> $book[[1]]$journal
#> [1] "Journal of Computational and Graphical Statistics 3:   :"
#> 
#> $book[[1]]$year
#> [1] "1996"
#> 
#> $book[[1]]$URL
#> [1] "https://doi.org/10.1080/10618600.1996.10474713"
#> 
#> $book[[1]]$isbn
#> NULL
#> 
#> $book[[1]]$pages
#> [1] "299--314"
#> 
#> 
#> $book[[2]]
#> $book[[2]]$unique_id
#> [1] "R"
#> 
#> $book[[2]]$author
#> [1] "R Core Team"
#> 
#> $book[[2]]$title
#> [1] "R: A Language and Environment for Statistical Computing"
#> 
#> $book[[2]]$journal
#> [1] "R Foundation for Statistical Computing Vienna Austria    :"
#> 
#> $book[[2]]$year
#> [1] "2016"
#> 
#> $book[[2]]$URL
#> [1] "https://www.R-project.org/"
#> 
#> $book[[2]]$isbn
#> [1] "3-900051-07-0"
#> 
#> $book[[2]]$pages
#> NULL
```

#### Stage 3: BibTeX writer

After reading the bibliographic entries and splitting out meaningful
values from them, we can finally write a structured file in the BibTeX
format.

The writer will read the bib chunks one at a time based on the metadata
extracted and will write the corresponding data fields. The default
entry type is a book, which should not have any problems with the web
articles.

``` r
file_path <- paste0(your_article_path,"/",file_name)
bib_path <- paste0(your_article_path,"/example.bib")
rebib:::bibtex_writer(bib_entry,file_path)
cat(readLines(paste(your_article_path,"example.bib",sep="/")),sep = "\n")
#> @book{ihaka:1996,
#> author = {{Ihaka, Ross and Gentleman, Robert}},
#> title = {{R: A Language for Data Analysis and Graphics}},
#> publisher = {Journal of Computational and Graphical Statistics 3:   :},
#> pages = {299--314},
#> year = {1996},
#> url = {https://doi.org/10.1080/10618600.1996.10474713}
#> }
#> @book{R,
#> author = {R {Core Team}},
#> title = {{R: A Language and Environment for Statistical Computing}},
#> publisher = {R Foundation for Statistical Computing Vienna Austria    :},
#> year = {2016},
#> url = {https://www.R-project.org/},
#> isbn = {3-900051-07-0}
#> }
```

Authors who are converting a bibliography from `bbl` should check if
there are any errors or missing values.

## General usage

General use of the [**rebib**](https://CRAN.R-project.org/package=rebib)
package is quite easy, requiring only one function call.

#### Using with RJournal article

In bibliography construction mode,
[**rebib**](https://CRAN.R-project.org/package=rebib) will check for
existing BibTeX files and will convert the embedded bibliography only
when there are no linked BibTeX files in the RJ article.

``` r
%% typically rebib will search for this line in the RJ article, 
%% not the RJwrapper file
\bibliography{example}
```

If there is no linked BibTeX file,
[**rebib**](https://CRAN.R-project.org/package=rebib) will begin its
procedure of converting the embedded bibliography to BibTeX and then
link it with the article file as well.

``` r
# for files without BibTeX source
rebib::handle_bibliography(your_article_path)
#> Warning in get_bib_file(article_dir, file_name): No Bib files found !
#> BibTeX file does not exist
#> will parse for bibliography
#> bibtex file created
cat(readLines(paste(your_article_path,"example.bib",sep="/")),sep = "\n")
#> @book{ihaka:1996,
#> author = {{Ihaka, Ross and Gentleman, Robert}},
#> title = {{R: A Language for Data Analysis and Graphics}},
#> publisher = {Journal of Computational and Graphical Statistics 3:   :},
#> pages = {299--314},
#> year = {1996},
#> url = {https://doi.org/10.1080/10618600.1996.10474713}
#> }
#> @book{R,
#> author = {R {Core Team}},
#> title = {{R: A Language and Environment for Statistical Computing}},
#> publisher = {R Foundation for Statistical Computing Vienna Austria    :},
#> year = {2016},
#> url = {https://www.R-project.org/},
#> isbn = {3-900051-07-0}
#> }
```

#### Bibliography aggregation

If you have a BibTeX file and it is missing some references then this
mode can help you read the embedded bibliography and combine the two
bibliographies in a single BibTeX file.

``` r
# for files with BibTeX source as well as embedded entries
rebib::aggregate_bibliography(your_article_path)
#> Found Bib file  example.bib
#> bibliography aggregation possible
#> Found Bib file  example.bib
#> aggregation delta : 2
#> BibTeX file aggregated
```

#### Using with any LaTeX/.bbl file

The bibliography converter will convert any LaTeX or `.bbl` file given
as a path. This way, one can generate BibTeX from the embedded
bibliography or `.bbl` file.

``` r
rebib::biblio_converter(path_to_your_file)
#> working directory : /temp/Rtmpt6Qk0v/exampledir/standalone
#> bib entries : 6
#> Written BibTeX file : /temp/Rtmpt6Qk0v/exampledir/standalone/sample.bib
```

## Limitations

The limitations of [**rebib**](https://CRAN.R-project.org/package=rebib)
package are:

-   The [**rebib**](https://CRAN.R-project.org/package=rebib) package
    only supports embedded bibliography format suggested/ commonly used
    in R Journal.

-   Encoding and character set issues might not work with rebib well.
    Try to use UTF-8 encoding if possible.

-   Bibliography aggregation won't fix existing issues like missing
    commas/brackets.

## Summary

In summary the [**rebib**](https://CRAN.R-project.org/package=rebib)
package supports:

-   Conversion of embedded natbib type bibliography entries used in the
    R Journal

-   Aggregation of bibliographic entries from the article as well as
    separate BibTeX files
:::