### 2025-09-07 supplement/metadata/RJwrapper.Rmd

---
title: Metadata and Other Structures in texor
abstract: |
  This is a small sample article to demonstrate handling of metadata and
  other structures during conversion in texor package.
author:
- name: Abhishek Ulayil
  affiliation: Institute of Actuaries of India (Student)
  orcid: 0009-0000-6935-8690
  address: Mumbai, India
date: '2025-09-01'
date_received: ~
journal:
  firstpage: ~
  lastpage: ~
volume: 17
issue: 3
slug: RJ-2025-000
citation_url: https://rjournal.github.io/
packages:
  cran: texor
  bioc: Biobase
preview: preview.png
bibliography: metadata.bib
CTV: ~
legacy_pdf: yes
legacy_converted: yes
output:
  rjtools::rjournal_web_article:
    self_contained: yes
    toc: no
    mathjax: https://cdn.jsdelivr.net/npm/mathjax@3/es5/tex-chtml-full.js
    md_extension: -tex_math_single_backslash

---

:::::::::: article
## Introduction

There is a difference in the structure of metadata and generally R
Markdown has more metadata entries than its LaTeX counterpart. So for
the conversion, the main task is to transform the metadata and include
it in the generated R Markdown file, matching the LaTeX metadata as
closely as possible.

When used with the `-s` or `standalone` mode, pandoc reads out some of
the existing meta data. The extracted metadata are listed below.

#### Metadata read from LaTeX file

-   Author names

-   Author address (affiliation)

-   ORCID number

-   Title

-   Bibliography

-   Abstract

-   CRAN/Bioconductor packages

The `\address{..}` section must be typeset in this manner for the most
effective conversion.

``` r
\address{
Abhishek Ulayil\\ % Your Name in the first Line
Institute of Actuaries of India (Student)\\% Your affiliation which includes post and/or org
Mumbai, India\\ % Your City/Country
ORCiD: 0009-0000-6935-8690\\% ORCID number with prefix ORCiD: 
}
```

#### Metadata read from DESCRIPTION file

-   Date

-   Slug

-   Volume and issue

Some metadata like the volume and issue were picked up from the folder
structure for some legacy articles, where the DESCRIPTION file was not
available.

#### Other supported commands in conversion

::: {#table:1}
  --------------------------------------------------------------------------------------------------------------------
             Command                                                    Render
  ------------------------------ -------------------------------------------------------------------------------------
          `\acronym{A}`                                                    A

               `\R`                                                        R

          `\pkg{texor}`                                                **texor**

        `\CRANpkg{texor}`                        [**texor**](https://CRAN.R-project.org/package=texor)

        `\BIOpkg{Biobase}`        [**Biobase**](https://www.bioconductor.org/packages/release/bioc/html/Biobase.html)

   `\ctv{ReproducibleResearch}`     [*ReproducibleResearch*](https://CRAN.R-project.org/view=ReproducibleResearch)

         `\command{help}`                                               `help`
  --------------------------------------------------------------------------------------------------------------------

  : (#tab:T1) Other commands supported by
  [**texor**](https://CRAN.R-project.org/package=texor) and pandoc
:::

## Miscellaneous {#misc}

#### Citations

Table [2](#tab:T2){reference-type="ref" reference="table:2"} represents
all the ways one can use citations in LaTeX and expect the generated
output.

::: {#table:2}
  ------------------------------------
          Command            Render
  ----------------------- ------------
     `\citet{pandoc}`       @pandoc

     `\citep{pandoc}`      [@pandoc]

     `\citet*{pandoc}`      @pandoc

     `\citep*{pandoc}`     [@pandoc]

    `\citeyear{pandoc}`    [-@pandoc]

   `\citeauthor{pandoc}`    @pandoc

      `\cite{pandoc}`      [@pandoc]
  ------------------------------------

  : (#tab:T2) Different forms of citation in LaTeX.
:::

#### Footnotes

Footnotes do not work in `tabular` environments in LaTeX but do work in
R Markdown tables[^1]. Custom footnote numbering works in LaTeX but does
not work in R Markdown[^2]. These features are demonstrated in
Table [3](#tab:T3){reference-type="ref" reference="table:3"}.

::: {#table:3}
  -----------------------------------------------------------------------
                             Command                              Render
  -------------------------------------------------------------- --------
   `\acronym{A}\footnote{This works in HTML, but not in LaTeX}`   A[^3]

          `\R\footnote[25]{HTML shows 4 rather than 25}`          R[^4]
  -----------------------------------------------------------------------

  : (#tab:T3) Demonstration of footnotes.
:::

#### Block quotes

> This block quote should justify an example.

#### Links

Most links work just fine using `\href` and the `\url` command will work
flawlessly, as demonstrated in Table  [4](#tab:T4){reference-type="ref"
reference="table:4"}.

::: {#table:4}
  ------------------------------------------------------------------------------------------------
               Command                                          Render
  --------------------------------- --------------------------------------------------------------
   `\href{www.google.com}{Google}`                 [Google](http://www.google.com)

       `\url{www.google.com}`               [www.google.com](http://www.google.com){.uri}

        `Table \ref{table:3}`        Table [3](#tab:T3){reference-type="ref" reference="table:3"}

      `Section \nameref{misc}`        Section [2](#misc){reference-type="ref" reference="misc"}

         `\autoref{table:3}`            [3](#tab:T3){reference-type="ref" reference="table:3"}
  ------------------------------------------------------------------------------------------------

  : (#tab:T4) Links
:::

##### Note

Certain reference commands like `\nameref{}` and `\autoref{}` may not
work as intended in HTML.

#### Lists

An example of usage of lists in LaTeX and **texor** generated HTML.

##### itemize

-   A

-   B

##### enumerate

1.  A

2.  B

##### description

**rebib**

:   Convert and Aggregate Bibliographies

**texor**

:   Converting 'LaTeX' 'R Journal' Articles into 'RJ-web-articles'

##### Note:

Some options for lists such as special numbering schemes or markups to
description item points might not work as intended and the generated
HTML might be simple.

#### Colored text

Some commands to describe colored text also work during conversion, such
as `\textcolor{blue}{Blue Text}` [Blue Text]{style="color: blue"}.

## Theorem environments

::: {#theorem:1 .theorem}
**Theorem 1**. *This is a sample theorem..*
:::

::: {#prop:1 .proposition}
**Proposition 1**. *As with $X \in \mathbb{R}$ this proposition \...*
:::

::: {#rm:1 .remark}
**Remark 1**. *[**texor**](https://CRAN.R-project.org/package=texor)
makes the article conversion easier.*
:::

**Pandoc** supports default LaTeX theorem environments. One can refer to
these environments like Theorem [1](#theorem:1){reference-type="ref"
reference="theorem:1"},Proposition [1](#prop:1){reference-type="ref"
reference="prop:1"} and Remark [1](#rm:1){reference-type="ref"
reference="rm:1"} made here. However custom theorem environments are not
supported yet.

## Summary

In summary the [**texor**](https://CRAN.R-project.org/package=texor)
package alongside pandoc supports:

-   Extracting metadata and putting it together well.

-   Handling links, block quotes, citations and footnotes.

-   Default theorem environments.
::::::::::

[^1]: Footnotes in paragraph text work in both LaTeX and R markdown

[^2]: Footnote 25

[^3]: This works in HTML, but not in LaTeX

[^4]: HTML shows 4 rather than 25